Thesis Title

### 2022-06-09 chapters/chapter_5.tex
Updates from Overleaf
Add introduction to chapter 6

A collection of de novo assembly tools was compiled, including \ac{OLC} and \ac{dBg} assembly algorithms, the latter including both single k-mer and multiple k-mer value approaches, and hybrid assemblers implementing both algorithms, including both genomic and metagenomic assemblers (Supplemental Table \ref{tab:ch5_suptable1}). Of these, 12 assemblers were selected based on the date of last update and are implemented in LMAS: ABySS \cite{jackman_abyss_2017} (version 2.3.1), BCALM2 \cite{chikhi_compacting_2016} (version 2.2.3), GATB Minia Pipeline \cite{noauthor_gatbgatb-minia-pipeline_2022} (commit hash 9d56f42) , IDBA-UD \cite{peng_idba-ud_2012} (version 1.1.3), MEGAHIT \cite{li_megahit_2015} (version 1.2.9), MetaHipMer2 \cite{georganas_extreme_2018} (version 2.0.0.65-gaad446d-dirty-AddGtest), metaSPAdes \cite{nurk_metaspades_2017} (version 3.15.3), minia \cite{chikhi_space-efficient_2013} (version 3.2.6), SKESA \cite{souvorov_skesa_2018} (version 2.5.0), SPAdes \cite{bankevich_spades_2012} (version 3.15.3), Unicycler \cite{wick_unicycler_2017} (version 0.4.9) and VelvetOptimiser \cite{seemann_velvetoptimiser_2021} (commit hash 092bdee) (Table \ref{tab:ch5_table1}). The execution commands for each assembler are available as Supplemental Material (see Supplemental Material, Short-read de novo assemblers) and in the documentation \cite{noauthor_short-read_nodate}. 
New assemblers can be added with minimal changes to the pipeline so that LMAS can be expanded as novel algorithms are developed. A template is available to facilitate their integration and a step-by-step guide is included in the documentation \cite{noauthor_add_nodate}. The only two requirements for the addition of a new assembler are the execution command for the assembler for paired-end short-read data and a Nextflow-compatible container with the assembler and any dependencies.

@@

The success of an assembly is evaluated in two steps: globally (see \ref{sssec:_chap5_global_metrics}) and relative to each of the replicons present in the sample (see \ref{sssec:_chap5_reference_metrics}). In both, the tabular presentation in the reports allows the comparison of exact values between assemblers, and the interactive plots allow a more intuitive overview and easy exploration of results. In addition to the assembly success metrics, computational resource statistics are registered for each assembler (see Supplemental Material, LMAS Metrics, Computational Performance Metrics).

@@

The computation of the global metrics is performed through statistics inherent to the complete set of contigs assembled per sample, independent of the species/sample of origin. The metrics are presented, in tabular form, for the complete set of contigs and those filtered for a minimum length, and also graphically for the contigs filtered for a minimum length. The statistics include information on contig number, size and ambiguous bases; and the proportion of reads mapping to the created contigs. Two statistics are a consolidation of per reference metrics: misassemblies (i.e. contigs that do not reflect the structural organisation in the reference replicons); and the overall size of gaps in all reference replicons not covered by any contig. A more detailed description of all global metrics is available in Supplemental Material (see Supplemental Material, LMAS Metrics, Global Metrics).

@@

For the computation of the reference-based metrics, only the \ac{FS} contigs are considered, for each reference replicon in the sample. These contigs are the ones exceeding the user-defined minimum sequence length, filtered using BBTools (version 38.44). After this initial step, the contigs are mapped to the reference replicons with minimap2 \cite{li_minimap2_2018} (version 2.22). The metrics are computed through custom python code (see Supplemental Material, Assembly filtering and mapping) for each replicon in the file provided as input. A detailed description of all reference-based metrics is available in Supplemental Material (see Supplemental Material, LMAS Metrics, Per Reference Metrics).

@@

The assessment and evaluation of genome assemblies has been a relevant field ever since the emergence of the assembly process itself, and therefore many solutions have been proposed \cite{sczyrba_critical_2017, olson_metagenomic_2019, bradnam_assemblathon_2013, gurevich_quast_2013, mikheenko_metaquast_2016, manchanda_genomeqc_2020, meader_genome_2010, challis_blobtoolkit_2020}. The Critical Assessment of Metagenome Interpretation (CAMI) proposed a set of recommendations and best practices for benchmarking in microbiome research \cite{meyer_tutorial_2021}. These recommendations include the reporting of computational performance, which may condition the choice of software by the users, such as runtime, disk space and memory consumption, also reported by  LMAS (see Supplementary Material, LMAS Metrics). As also suggested by CAMI, LMAS tracks the exact program version and command-line calls through its implementation in Nextflow. Moreover,  using containerised assemblers and being easily installable through Bioconda, LMAS facilitates deployment in diverse user machines. Unlike the CAMI tutorial, in which users are asked to download and install the necessary tools, in LMAS everything is provided in a one-stop reproducible workflow that effortlessly handles all pre-processing, assembly, post-processing, traceability and report production steps, freeing users to focus on providing relevant samples for analysis and interpreting the results in view of the intended applications.

Concerning software for assembly quality assessment currently available, the most widely adopted is QUAST \cite{gurevich_quast_2013}, or when dealing with metagenomic data, its extension metaQUAST \cite{mikheenko_metaquast_2016}, which was also adopted by the CAMI challenges [3,5] \cite{sczyrba_critical_2017, meyer_critical_2021} and suggested in the CAMI Tutorial \cite{meyer_tutorial_2021}. Although several features of these tools overlap with LMAS’ quality assessment components, these differ from LMAS in the sense that they are not a single step workflow allowing a traceable and reproducible assembly of mock communities. Unlike QUAST and metaQUAST, whose purpose is to evaluate assemblies, the purpose of LMAS is to allow users to evaluate assembler performance for a given sample of interest. Supplementary Table \ref{tab:ch5_suptable} shows the comparison of the output and computed assembly quality metrics generated by LMAS, QUAST and metaQUAST.

@@

To illustrate the use of LMAS and evaluate the performance of the chosen assemblers we used the eight bacterial genomes and four plasmids of the ZymoBIOMICS Microbial Community Standards as reference. As input we used the raw sequence reads of mock communities with an even and logarithmic distribution of species, from real sequencing runs \cite{nicholls_ultra-deep_2019} and simulated read datasets, with and without error, matching the distribution of species in each sample \cite{gourle_simulating_2019}. Our dataset is composed of samples ENN (in silico generated evenly distributed without error), EMS (in silico generated evenly distributed with Illumina MiSeq error model), ERR2984773 (evenly distributed real Illumina MiSeq sample), LNN (in silico generated logarithmically distributed without error), LHS (in silico generated logarithmically distributed with Illumina HiSeq error model) and ERR2935805 (logarithmically distributed real Illumina HiSeq sample) (see Supplemental Table \ref{tab:ch5_suptable3}). Detailed information about the generation of the input samples is available as Supplemental Material (see Supplemental Materials, ZymoBIOMICS microbial community standards, Supplemental Table \ref{tab:ch5_suptable4}). To evaluate the reproducibility of an assembler performance, the LMAS workflow was run three times for all samples using default parameters, and the resulting data was processed for each sample (see Supplemental Materials, Assessment of Assembly Success) Supplementary Table \ref{tab:ch5_suptable5} to Table \ref{tab:ch5_suptable10} present an overview of the average global performance per assembler for each sample in LMAS.

@@

Of the 12 de novo prokaryotic assemblers included in LMAS, five stand out as having an overall poor performance: ABySS, BCALM2, MetaHipmer2, minia and VelvetOptimiser. Both ABySS and MetaHipmer2 performed inconsistently with differing resource requirements for the same sample in different runs, namely run time and memory allocation (see Supplemental Materials, Resource Requirements Differ Greatly, Supplemental Figure \ref{fig:chap5_sup_figure_2}).  Moreover, ABySS failed to produce an assembly for sample ERR2984773 for 1 of the runs (see Supplementary Table \ref{tab:ch5_suptable7}) and for sample LHS in any of the 3 runs in the time limit of 3 days (see Supplementary Table \ref{tab:ch5_suptable9}), and MetaHipmer2 failed to produce an assembly for samples LNN and LHS in all 3 runs (see Supplementary Tables \ref{tab:ch5_suptable8}-\ref{tab:ch5_suptable9}). VelvetOptimiser generated the highest number of inconsistent contigs across the 3 LMAS runs (Figure \ref{fig:chap5_figure3}, Supplementary Table \ref{tab:ch5_suptable11}), with 1.69\% of the total contigs created present in only 1 or 2 runs. Although not as extreme as VelvetOptimiser, ABySS (0.52\%), minia (0.14\%), GATBMiniaPipeline (0.32\%), MetaHipMer2 (0.11\%) and IDBA-UD (0.08\%) also showed inconsistencies in contig size.

@@

Regarding the quality assessment of the assemblies produced (Figure \ref{fig:chap5_figure4}, Supplementary Table \ref{tab:ch5_suptable12}), ABySS, BCALM2 and minia are the only single k-mer \ac{dBg} assemblers in the collection and were found to mostly underperform relative to their multiple k-mer \ac{dBg} counterparts, generally resulting in more fragmented assemblies, although there were significant differences in performance across samples. Among multiple k-mer assemblers, VelvetOptimiser frequently produced a very high number of contigs of very small size (over 99\% of the contigs not surpassing the minimum length of 1,000 \ac{bp}) and therefore a low N50 (an average of 29,768 \ac{bp} versus a global average of 84,114 \ac{bp}) (Supplementary tables \ref{tab:ch5_suptable5}-\ref{tab:ch5_suptable10}). Additionally, ABySS and VelvetOptimizer produced contigs with a very large number of Ns, with an average of 1,019 and 3,035 uncalled bases per assembly, respectively. MetaHipMer2, although having overall average metrics in the two evenly distributed mock samples (ENN and EMS, Supplementary Tables \ref{tab:ch5_suptable5}-\ref{tab:ch5_suptable6}) where it was able to run successfully, it severely underperformed in the real samples (ERR2984773 and ERR2935805, Supplementary Tables \ref{tab:ch5_suptable7} and \ref{tab:ch5_suptable10}). Generally, the performance scores of the assemblers decreased considerably for the real samples in comparison with the simulated ones, either with or without error. High utilisation of the reads in the dataset is observed for most assemblers, with on average at least 90\% of the reads mapping back to the assembly, except for ABySS, BCALM2, MetaHipMer2 and VelvetOptimiser whose values are in the range of 46-79\%. Despite an overall good performance, SPAdes produced the highest number of misassembled contigs, with an average of 98 and a maximum of 572 (sample ERR2935805, Supplementary \ac{Pls}), in comparison to the global average of 11 misassembled contigs for all assemblers across all samples.

@@

\subsection{Short-read de novo assemblers}

@@

\subsection{Assembly filtering and mapping}

@@

\subsection{LMAS Metrics}

@@

\subsubsection{Global Metrics}

@@

\subsubsection{Per Reference Metrics}

@@

\subsubsection{Computational Performance Metrics}

@@

\subsection{ZymoBIOMICS microbial community standards}

@@

\subsubsection{Assessment of Assembly Success}

@@

\subsubsection{Resource requirements differ greatly}

@@

\input{tables/chapter_5/suptable1}
\input{tables/chapter_5/suptable2}
\input{tables/chapter_5/suptable3}
\input{tables/chapter_5/suptable4}
\input{tables/chapter_5/suptable5}
\input{tables/chapter_5/suptable6}
\input{tables/chapter_5/suptable7}
\input{tables/chapter_5/suptable8}
\input{tables/chapter_5/suptable9}
\input{tables/chapter_5/suptable10}
\input{tables/chapter_5/suptable11}
\input{tables/chapter_5/suptable12}
\input{tables/chapter_5/suptable13}
\input{tables/chapter_5/suptable14}

>>>>

A collection of de novo assembly tools was compiled, including \ac{OLC} and \ac{dBg} assembly algorithms, the latter including both single k-mer and multiple k-mer value approaches, and hybrid assemblers implementing both algorithms, including both genomic and metagenomic assemblers (Supplemental Table \ref{tab:ch5_suptable1}). Of these, 12 assemblers were selected based on the date of last update and are implemented in LMAS: ABySS \cite{jackman_abyss_2017} (version 2.3.1), BCALM2 \cite{chikhi_compacting_2016} (version 2.2.3), GATB Minia Pipeline \cite{noauthor_gatbgatb-minia-pipeline_2022} (commit hash 9d56f42) , IDBA-UD \cite{peng_idba-ud_2012} (version 1.1.3), MEGAHIT \cite{li_megahit_2015} (version 1.2.9), MetaHipMer2 \cite{georganas_extreme_2018} (version 2.0.0.65-gaad446d-dirty-AddGtest), metaSPAdes \cite{nurk_metaspades_2017} (version 3.15.3), minia \cite{chikhi_space-efficient_2013} (version 3.2.6), SKESA \cite{souvorov_skesa_2018} (version 2.5.0), SPAdes \cite{bankevich_spades_2012} (version 3.15.3), Unicycler \cite{wick_unicycler_2017} (version 0.4.9) and VelvetOptimiser \cite{seemann_velvetoptimiser_2021} (commit hash 092bdee) (Table \ref{tab:ch5_table1}). The execution commands for each assembler are available as Supplemental Material (see \ref{ch5_sup_assemblers}) and in the documentation \cite{noauthor_short-read_nodate}. 
New assemblers can be added with minimal changes to the pipeline so that LMAS can be expanded as novel algorithms are developed. A template is available to facilitate their integration and a step-by-step guide is included in the documentation \cite{noauthor_add_nodate}. The only two requirements for the addition of a new assembler are the execution command for the assembler for paired-end short-read data and a Nextflow-compatible container with the assembler and any dependencies.

@@

The success of an assembly is evaluated in two steps: globally (see \ref{sssec:_chap5_global_metrics}) and relative to each of the replicons present in the sample (see \ref{sssec:_chap5_reference_metrics}). In both, the tabular presentation in the reports allows the comparison of exact values between assemblers, and the interactive plots allow a more intuitive overview and easy exploration of results. In addition to the assembly success metrics, computational resource statistics are registered for each assembler (see \ref{ch5_sup_performance_metrics}).

@@

The computation of the global metrics is performed through statistics inherent to the complete set of contigs assembled per sample, independent of the species/sample of origin. The metrics are presented, in tabular form, for the complete set of contigs and those filtered for a minimum length, and also graphically for the contigs filtered for a minimum length. The statistics include information on contig number, size and ambiguous bases; and the proportion of reads mapping to the created contigs. Two statistics are a consolidation of per reference metrics: misassemblies (i.e. contigs that do not reflect the structural organisation in the reference replicons); and the overall size of gaps in all reference replicons not covered by any contig. A more detailed description of all global metrics is available in Supplemental Material (see \ref{ch5_supmaterial_metrics_global}).

@@

For the computation of the reference-based metrics, only the \ac{FS} contigs are considered, for each reference replicon in the sample. These contigs are the ones exceeding the user-defined minimum sequence length, filtered using BBTools (version 38.44). After this initial step, the contigs are mapped to the reference replicons with minimap2 \cite{li_minimap2_2018} (version 2.22). The metrics are computed through custom python code (see \ref{ch5_sup_assembly_filtering_and_mapping}) for each replicon in the file provided as input. A detailed description of all reference-based metrics is available in Supplemental Material (see \ref{ch5_supmaterial_metrics_reference}).

@@

The assessment and evaluation of genome assemblies has been a relevant field ever since the emergence of the assembly process itself, and therefore many solutions have been proposed \cite{sczyrba_critical_2017, olson_metagenomic_2019, bradnam_assemblathon_2013, gurevich_quast_2013, mikheenko_metaquast_2016, manchanda_genomeqc_2020, meader_genome_2010, challis_blobtoolkit_2020}. The Critical Assessment of Metagenome Interpretation (CAMI) proposed a set of recommendations and best practices for benchmarking in microbiome research \cite{meyer_tutorial_2021}. These recommendations include the reporting of computational performance, which may condition the choice of software by the users, such as runtime, disk space and memory consumption, also reported by  LMAS (see \ref{ch5_supmaterial_metrics}). As also suggested by CAMI, LMAS tracks the exact program version and command-line calls through its implementation in Nextflow. Moreover,  using containerised assemblers and being easily installable through Bioconda, LMAS facilitates deployment in diverse user machines. Unlike the CAMI tutorial, in which users are asked to download and install the necessary tools, in LMAS everything is provided in a one-stop reproducible workflow that effortlessly handles all pre-processing, assembly, post-processing, traceability and report production steps, freeing users to focus on providing relevant samples for analysis and interpreting the results in view of the intended applications.

Concerning software for assembly quality assessment currently available, the most widely adopted is QUAST \cite{gurevich_quast_2013}, or when dealing with metagenomic data, its extension metaQUAST \cite{mikheenko_metaquast_2016}, which was also adopted by the CAMI challenges [3,5] \cite{sczyrba_critical_2017, meyer_critical_2021} and suggested in the CAMI Tutorial \cite{meyer_tutorial_2021}. Although several features of these tools overlap with LMAS’ quality assessment components, these differ from LMAS in the sense that they are not a single step workflow allowing a traceable and reproducible assembly of mock communities. Unlike QUAST and metaQUAST, whose purpose is to evaluate assemblies, the purpose of LMAS is to allow users to evaluate assembler performance for a given sample of interest. Supplementary Table \ref{tab:ch5_suptable2} shows the comparison of the output and computed assembly quality metrics generated by LMAS, QUAST and metaQUAST.

@@

To illustrate the use of LMAS and evaluate the performance of the chosen assemblers we used the eight bacterial genomes and four plasmids of the ZymoBIOMICS Microbial Community Standards as reference. As input we used the raw sequence reads of mock communities with an even and logarithmic distribution of species, from real sequencing runs \cite{nicholls_ultra-deep_2019} and simulated read datasets, with and without error, matching the distribution of species in each sample \cite{gourle_simulating_2019}. Our dataset is composed of samples ENN (in silico generated evenly distributed without error), EMS (in silico generated evenly distributed with Illumina MiSeq error model), ERR2984773 (evenly distributed real Illumina MiSeq sample), LNN (in silico generated logarithmically distributed without error), LHS (in silico generated logarithmically distributed with Illumina HiSeq error model) and ERR2935805 (logarithmically distributed real Illumina HiSeq sample) (see Supplemental Table \ref{tab:ch5_suptable3}). Detailed information about the generation of the input samples is available as Supplemental Material (see \ref{ch5_sup_zymobiomics}, Supplemental Table \ref{tab:ch5_suptable4}). To evaluate the reproducibility of an assembler performance, the LMAS workflow was run three times for all samples using default parameters, and the resulting data was processed for each sample (see Chapter \ref{ch5_sup_assembly_success}) Supplementary Table \ref{tab:ch5_suptable5} to Table \ref{tab:ch5_suptable10} present an overview of the average global performance per assembler for each sample in LMAS.

@@

Of the 12 de novo prokaryotic assemblers included in LMAS, five stand out as having an overall poor performance: ABySS, BCALM2, MetaHipmer2, minia and VelvetOptimiser. Both ABySS and MetaHipmer2 performed inconsistently with differing resource requirements for the same sample in different runs, namely run time and memory allocation (see \ref{ch5_sup_resources}, Supplemental Figure \ref{fig:chap5_sup_figure_2}).  Moreover, ABySS failed to produce an assembly for sample ERR2984773 for 1 of the runs (see Supplementary Table \ref{tab:ch5_suptable7}) and for sample LHS in any of the 3 runs in the time limit of 3 days (see Supplementary Table \ref{tab:ch5_suptable9}), and MetaHipmer2 failed to produce an assembly for samples LNN and LHS in all 3 runs (see Supplementary Tables \ref{tab:ch5_suptable8}-\ref{tab:ch5_suptable9}). VelvetOptimiser generated the highest number of inconsistent contigs across the 3 LMAS runs (Figure \ref{fig:chap5_figure3}, Supplementary Table \ref{tab:ch5_suptable11}), with 1.69\% of the total contigs created present in only 1 or 2 runs. Although not as extreme as VelvetOptimiser, ABySS (0.52\%), minia (0.14\%), GATBMiniaPipeline (0.32\%), MetaHipMer2 (0.11\%) and IDBA-UD (0.08\%) also showed inconsistencies in contig size.

@@

Regarding the quality assessment of the assemblies produced (Figure \ref{fig:chap5_figure4}, Supplementary Table \ref{tab:ch5_suptable12}), ABySS, BCALM2 and minia are the only single k-mer \ac{dBg} assemblers in the collection and were found to mostly underperform relative to their multiple k-mer \ac{dBg} counterparts, generally resulting in more fragmented assemblies, although there were significant differences in performance across samples. Among multiple k-mer assemblers, VelvetOptimiser frequently produced a very high number of contigs of very small size (over 99\% of the contigs not surpassing the minimum length of 1,000 \ac{bp}) and therefore a low N50 (an average of 29,768 \ac{bp} versus a global average of 84,114 \ac{bp}) (Supplementary Tables \ref{tab:ch5_suptable5}-\ref{tab:ch5_suptable10}). Additionally, ABySS and VelvetOptimizer produced contigs with a very large number of Ns, with an average of 1,019 and 3,035 uncalled bases per assembly, respectively. MetaHipMer2, although having overall average metrics in the two evenly distributed mock samples (ENN and EMS, Supplementary Tables \ref{tab:ch5_suptable5}-\ref{tab:ch5_suptable6}) where it was able to run successfully, it severely underperformed in the real samples (ERR2984773 and ERR2935805, Supplementary Tables \ref{tab:ch5_suptable7} and \ref{tab:ch5_suptable10}). Generally, the performance scores of the assemblers decreased considerably for the real samples in comparison with the simulated ones, either with or without error. High utilisation of the reads in the dataset is observed for most assemblers, with on average at least 90\% of the reads mapping back to the assembly, except for ABySS, BCALM2, MetaHipMer2 and VelvetOptimiser whose values are in the range of 46-79\%. Despite an overall good performance, SPAdes produced the highest number of misassembled contigs, with an average of 98 and a maximum of 572 (sample ERR2935805, , Supplementary Table \ref{tab:ch5_suptable10}), in comparison to the global average of 11 misassembled contigs for all assemblers across all samples.

@@

\subsection{Short-read de novo assemblers} \label{ch5_sup_assemblers}

@@

\subsection{Assembly filtering and mapping} \label{ch5_sup_assembly_filtering_and_mapping}

@@

\subsection{LMAS Metrics} \label{ch5_supmaterial_metrics}

@@

\subsubsection{Global Metrics} \label{ch5_supmaterial_metrics_global}

@@

\subsubsection{Per Reference Metrics} \label{ch5_supmaterial_metrics_reference}

@@

\subsubsection{Computational Performance Metrics} \label{ch5_sup_performance_metrics}

@@

\subsection{ZymoBIOMICS microbial community standards} \label{ch5_sup_zymobiomics}

@@

\subsubsection{Assessment of Assembly Success} \label{ch5_sup_assembly_success}

@@

\subsubsection{Resource requirements differ greatly} \label{ch5_sup_resources}

@@

\input{tables/chapter_5/suptable1}
\input{tables/chapter_5/suptable2}
\input{tables/chapter_5/suptable3}
\input{tables/chapter_5/suptable4}
\input{tables/chapter_5/suptable5}
\input{tables/chapter_5/suptable6}
\input{tables/chapter_5/suptable7}
\input{tables/chapter_5/suptable8}
\input{tables/chapter_5/suptable9}
\input{tables/chapter_5/suptable10}
\input{tables/chapter_5/suptable11}
\input{tables/chapter_5/suptable12}
\input{tables/chapter_5/suptable13}
\input{tables/chapter_5/suptable14}
\input{tables/chapter_5/suptable15}
\input{tables/chapter_5/suptable16}
\input{tables/chapter_5/suptable17}
\input{tables/chapter_5/suptable18}
\input{tables/chapter_5/suptable19}
\input{tables/chapter_5/suptable20}
\input{tables/chapter_5/suptable21}
\input{tables/chapter_5/suptable22}
\input{tables/chapter_5/suptable23}

### 2022-06-09 chapters/chapter_6.tex
Add introduction to chapter 6

>>>>

As discussed in Chapters \ref{ch:introduction} \ref{ch:paper1} and \ref{ch:paper2}, \ac{SMg} can offer relatively unbiased pathogen detection and characterisation, potentially able to provide antimicrobial resistance and virulence profiling in a single methodological step. The rise of \ac{AMR} poses a major threat to human health around the world.

Despite the relative standardisation of the process to acquire genomic data, either through \ac{WGS} or \ac{SMg}, there's a myriad of different tools available to perform \textit{in silico} \ac{AMR} detection, with some using their own reference database and each one generating a unique, non-standardised report of the genes or variants that can possibly confer resistance in a given sample. This is a huge barrier to the comparison
of results and the modularity of tools within bioinformatic workflows.

In this chapter we present a a standardised output specification for the reporting of genes or variants potentially conferring \ac{AMR}. This specification aims to standardise the reporting of these results, regardless of the tool and/or database used. To ease its utilisation, we've packaged this specification into hAMRonization, a command-line utility that is able to aggregate results form a wide variety of \ac{AMR} detection tools, both species-agnostic and species-specific, providing a unified report tabular form, JSON or through an interactive HTML file that can be opened within the browser for navigable data exploration.

We aim that with hAMRonization, stakeholders, such as clinical practitioners, will be more efficiently informed of potential \ac{AMR} genes of variants in a given dataset containing genomic information. Ultimately, the possibility of using multiple \ac{AMR} detection tools, with multiple \ac{AMR} databases, increases evidence, which increases confidence in the results obtained. Likewise, differences in databases can be evidenced fir better interpretability.
